Overview

### 2023-12-07 analysis/001_overview.Rmd
Country plot top 20

>>>>

And here's my country analysis. First top 20

```{r coutnry_data_prep}
country_all <- hoaddata::jn_aff |>
  filter(!cr_year %in% c(2017, 2023), issn_l != "0027-8424") |>
  # only with cc
  filter(issn_l %in% hoaddata::cc_articles$issn_l) |>
  distinct(issn_l, cr_year, country_code, articles_total) |>
  group_by(cr_year, country_code) |>
  summarise(all = sum(articles_total))

country_oa <- hoaddata::cc_articles |>
  filter(!cr_year %in% c(2017, 2023), issn_l != "0027-8424") |>
  group_by(cr_year, country_code) |>
  summarise(oa = n_distinct(doi))

country_codes <- readr::read_csv(here::here("data", "ror_country_codes.csv"))
country_ta <-  ta_oa_inst |>
  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
  left_join(country_codes) |>
  group_by(cr_year, country_code) |>
  summarise(ta_oa = n_distinct(doi))

country_df <- country_all |>
  left_join(country_oa, by = c("cr_year", "country_code")) |>
  left_join(country_ta, by = c("cr_year", "country_code")) |>
  distinct() |>
  ungroup() |>
  mutate(cr_year = as.character(cr_year)) |>
  mutate(across(c(oa, ta_oa), .fns = ~replace_na(.,0))) |>
  filter(!is.na(country_code)) |>
  mutate(
    country_name = countrycode::countrycode(country_code, origin = "iso2c", destination = "country.name")
  )

# Top 20 in terms of articles published
top_20 <- country_df |>
  filter(!is.na(country_code)) |>
  group_by(country_code, country_name) |>
  summarise(n = sum(all)) |>
  arrange(desc(n)) |>
  head(20)

## Order of multiples, sort by the most productive (Top 20)

country_multiple_df <- country_df |>
  mutate(country_name_fct = forcats::fct_other(country_name, keep = top_20$country_name)) |>
  mutate(country_name_fct = forcats::fct_relevel(country_name_fct, top_20$country_name, "Other")) |>
  group_by(country_name_fct, cr_year) |>
  summarise(
    all_total = sum(all),
    oa_total = sum(oa),
    ta_oa_total = sum(ta_oa)
  ) |>
  mutate(prop = oa_total / all_total) |>
  # remove cases where we could not retrieve a country name |>
  filter(!is.na(country_name_fct))

# Prepare facet heading
facet_labels <- country_multiple_df |>
  group_by(country_name_fct) |>
  summarise(total = sum(all_total)) |>
  mutate(prop = total / sum(total)) |>
  ungroup() |>
  # Highlight Germany
  mutate(total_string = case_when(
    total < 1000 ~ as.character(total),
    total < 1000000 ~ paste0(format(round(total / 1e3, 1), trim = TRUE), "K"),
    total >= 1000000 ~ paste0(format(round(total / 1e6, 2), trim = TRUE), "M")
  )) |>
  mutate(
    facet_labels = glue::glue(
      '{country_name_fct}<br>
      {total_string} | {paste(round(prop * 100, 1), "%")}'
    )
  ) |>
  distinct(country_name_fct, facet_labels)

country_facet_labels <- as.character(facet_labels$facet_labels)
names(country_facet_labels) <- facet_labels$country_name_fct

top_20_plot_df <- 
  country_multiple_df |> 
  mutate(no_ta = oa_total - ta_oa_total) |>
  rename(ta_oa = ta_oa_total) |>
  pivot_longer(
  cols = c(no_ta, ta_oa),
  values_to = "value",
  names_to = "type") 
```

```{r, country_plot, fig.asp=0.8}
ggplot(top_20_plot_df, aes(
    gsub("^20", "'", as.character(cr_year)),
    value / all_total,
    group = type,
    fill = type
  )) +
  geom_area(stat = "identity") +
  scale_fill_manual(
    "Hybrid OA",
    values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"),
    labels = c(no_ta = "Other", ta_oa = "Agreement"),
    guide = guide_legend(reverse = TRUE)
  ) +
  facet_wrap( ~ country_name_fct,
              nrow = 3,
              labeller = as_labeller(country_facet_labels)) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1L),
    expand = expansion(mult = c(0, 0.05)),
    breaks = scales::breaks_extended(4)
  ) +
  theme_minimal() +
  labs(y = "OA Share", x = "2018 - 2022") +
  theme(
    legend.position = "top",
    legend.justification = "right",
    legend.direction = "horizontal",
    panel.grid.minor = element_blank(),
                panel.border = element_rect(color = "grey50", fill = NA),
    axis.title=element_text(size = 10),
    strip.text = ggtext::element_markdown(size = 6),
    plot.subtitle = ggtext::element_markdown(hjust = 0.5),
    panel.spacing = unit(1, "lines")
  )
```
